---
title: "Project 6"
output: html_document
---
## Certificate
[Certificate](assets/CertificateOfCompletion.pdf)

## Summary of Concepts Learned
In Prof. Paulsons course going over R for Data Science: Analysis and Visualization I learned several different useful applications for R Studio, such as:
- Data Wrangling to filter, sort, change, and summarize datasets
- Data Visualization with various graphs, charts, and histograms
- Data Importation using CSV from external sources to analyzing it within R Studio
- Replicating data through R Notebooks and RMarkdown
- How to tidy datasets, utilizing 'tidyr'
- How to explore real world datasets using RStudio

## Finding statistics of a dataset using internal parameters 
```{r}
library(tidyverse)

data("mtcars")

# groups data by cylinders, find mpg/weight
mtcars_summary <- mtcars %>%
  group_by(cyl) %>%
  summarise(
    avg_mpg = mean(mpg),
    avg_wt = mean(wt)
  )

print(mtcars_summary)
```

## Using ChatGPT to determine the spread of EV registration across the United States
```{r}
library(tidyverse)
ev_data <- read_csv("EV_Cleaned.csv")


# Get the top 3 states by registration count
top3 <- ev_data %>%
  arrange(desc(`Registration Count`)) %>%
  slice(1:3)

# Calculate total national registrations
total_national <- sum(ev_data$`Registration Count`)

# Calculate 'Other States' total
others_total <- total_national - sum(top3$`Registration Count`)

# Create final data frame for plotting
pie_data <- top3 %>%
  select(State, `Registration Count`) %>%
  add_row(State = "Other States", `Registration Count` = others_total)

# Create pie chart
pie(pie_data$`Registration Count`,
    labels = paste0(pie_data$State, " (", round(100 * pie_data$`Registration Count` / sum(pie_data$`Registration Count`), 1), "%)"),
    main = "Top 3 States vs Rest of US - EV Registrations (2023)",
    col = rainbow(4))

```
